Distributional Trees and Forests

### 2017-10-08 pkg/disttree/inst/draft/draft.Rnw
restructured plots
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1635 255f4249-2263-4417-b1ea-8498779e27a7

    # add labels
    text(x = -0.8, y = lh1, labels = "2010", col = "red", cex = 0.8)
    text(x = -0.8, y = lh2, labels = "2011", col = "blue", cex = 0.8)
    text(x = -0.8, y = lh3, labels = "2012", col = "darkgreen", cex = 0.8)
@
\caption{\label{}}
\end{figure}

@@

\begin{figure}[t!]
<<fig = TRUE, echo = FALSE>>=
par(mfrow = c(1,2))
plot(x = sp1[,id_mi[1]], y = sp1[, "mu"], type = "l", xlab = names_mi[1], ylab = "mu")
plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = "sigma")
@
\end{figure}
marginal effect of covariate with highest variable importance in distforest\\
(other covariates? no effects for 4th variable, probably because only splits after several other splits)

@@

\subsubsection{Cross-Validation}
randomly split into 10 (roughly) equally sized data sets, learn on 9, evaluate on the 10th\\
average over 10 variations\\
repeat 10 times\\
compare rmse, ll, crps score skills with reference model EMOS log\\

@@

\begin{figure}[t!]
\begin{subfigure}{0.3\textwidth}
<<rain_cross_axams_rmse, fig=TRUE, echo=FALSE>>=
boxplot(rain_rmse)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_cross_axams_ll, fig=TRUE, echo=FALSE>>=
boxplot(rain_ll, ylim = c(-2.8,-1.3))
#boxplot(rain_ll)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_cross_axams_crps, fig=TRUE, echo=FALSE>>=
boxplot(rain_crps)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_cross_axams_crps_skills_score, fig=TRUE, echo=FALSE>>=
boxplot(1 - rain_crps / rain_crps[,6]) 
abline(h = 0, col = 2, lwd = 2)
@
\end{subfigure}
\caption{\label{boxplots_rain}Results 10x10 cross validation at the station Axams (RMSE, log-likelihood and CRPS)\\
left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference between the observations and the estimated expected value\\
right: continuous lines represent the location parameter, dotted lines the scale parameter}
\end{figure}

\subsection{Predictions over all stations}

NAs in gamlss in 13 stations!!!\\
% sum(is.na(crps[,3]))
<<eval = TRUE, echo = FALSE>>=
#### prediction over all stations 25 - 3
if(file.exists("rain_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred.rda")
  load("rain_pred.rda")
} else {

@@

\begin{figure}[t!]
\begin{subfigure}{0.3\textwidth}
<<rain_pred_rmse, fig=TRUE, echo=FALSE>>=
boxplot(rmse)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_pred_ll, fig=TRUE, echo=FALSE>>=
boxplot(ll, ylim = c(-2.8,-1.3))
#boxplot(ll)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_pred_crps, fig=TRUE, echo=FALSE>>=
boxplot(crps)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.3\textwidth}
<<rain_pred_skills_score, fig=TRUE, echo=FALSE>>=
# skills score
  boxplot(1 - crps/crps[,6])
  abline(h = 0, col = "red")
@
\end{subfigure}
\caption{\label{boxplots_rain}Results of predictions over all stations (RMSE, log-likelihood and CRPS)\\
left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference between the observations and the estimated expected value\\
right: continuous lines represent the location parameter, dotted lines the scale parameter}
\end{figure}

>>>>

    # add labels
    text(x = -0.8, y = lh1, labels = "2010", col = "red", cex = 0.8)
    text(x = -0.8, y = lh2, labels = "2011", col = "blue", cex = 0.8)
    text(x = -0.8, y = lh3, labels = "2012", col = "darkgreen", cex = 0.8)
@
\caption{\label{}Predictions at station Axams for one day in July of 2010, 2011 and 2012}
\end{figure}

@@

\begin{figure}[t!]
\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
plot(x = sp1[,id_mi[1]], y = sp1[, "mu"], type = "l", xlab = names_mi[1], ylab = "mu")
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = "sigma")
@
\end{subfigure}
\caption{\label{varimp}marginal effect of covariate with highest variable importance in distforest}
\end{figure}
(other covariates? no effects for 4th variable, probably because only splits after several other splits)

@@

\newpage
\subsubsection{Cross-Validation}
randomly split into 10 (roughly) equally sized data sets, learn on 9, evaluate on the 10th\\
average over 10 variations\\
repeat 10 times\\
compare rmse, ll, crps score skills with reference model EMOS log\\

@@

\begin{figure}[t!]
\begin{subfigure}{0.5\textwidth}
<<rain_cross_axams_rmse, fig=TRUE, echo=FALSE>>=
boxplot(rain_rmse)
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.5\textwidth}
<<rain_cross_axams_ll, fig=TRUE, echo=FALSE>>=
boxplot(rain_ll, ylim = c(-2.8,-1.3))
#boxplot(rain_ll)
@
\end{subfigure}
\caption{\label{boxplots_rain}Results 10x10 cross validation at the station Axams\\
left: RMSE, right: log-likelihood}
\end{figure}

\begin{figure}[t!]
<<rain_cross_axams_crps_skills_score, fig=TRUE, echo=FALSE>>=
boxplot(1 - rain_crps / rain_crps[,6]) 
abline(h = 0, col = 2, lwd = 2)
@
\caption{\label{boxplots_rain}Results 10x10 cross validation at the station Axams \\
CRPS skills score with reference model EMOS log}
\end{figure}

\newpage
\subsection{Predictions over all stations}

NAs in gamlss in 13 stations!!!\\
for example in 13th station in complete{\_}stations\\
\text{Error in glim.fit(f = mu.object, X = mu.X, y = y, w = w, fv = mu, os = mu.offset,  : 
  NA's in the working vector or weights for parameter mu}
\\
% sum(is.na(crps[,3]))
<<eval = TRUE, echo = FALSE>>=
#### prediction over all stations 25 - 3
if(file.exists("rain_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred.rda")
  load("rain_pred.rda")
} else {

@@

\begin{figure}[t!]
<<rain_pred_skills_score, fig=TRUE, echo=FALSE>>=
# skills score
  boxplot(1 - crps/crps[,6])
  abline(h = 0, col = "red")
@
\caption{\label{boxplots_rain}Results of predictions over all stations: CRPS skills score with reference model EMOS log}
\end{figure}
